---
title: "Abiotic data"
author: "Chiara Duijser"
date: "2023-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r}
library(plyr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(RColorBrewer)
```

## Load in data for the Inner Mangrove logger
```{r}
m_log <- read.csv("../2022-2023 Mangrove HOBO_Temp_pH.csv", header=FALSE, comment.char="#", stringsAsFactors=TRUE)

# The time variable isn't at the date format, so time series data won't work
m_log$Date_Time <- as.POSIXct(m_log$Date_Time, format = "%d/%m/%Y %H:%M") 

# Set the range of datetimes (1 year). First few days are omitted from the dataset as logger needs to equilibrate first few days.
m_start_time <- as.POSIXct("2022-02-10 0:16:20", format = "%Y-%m-%d %H:%M:%S")
m_end_time <- as.POSIXct("2023-02-10 23:16:20", format = "%Y-%m-%d %H:%M:%S")

# Subset the data within the datetime range 
m_subset <- m_log[m_log$Date_Time >= m_start_time & m_log$Date_Time <= m_end_time, ]

# Calculate summary statistics for temperature
m_mean_temp <- mean(m_subset$Temperature, na.rm=TRUE)
m_max_temp <- max(m_subset$Temperature, na.rm=TRUE)
m_min_temp <- min(m_subset$Temperature, na.rm=TRUE)
m_sd_temp <- sd(m_subset$Temperature, na.rm=TRUE)
m_se_temp <- m_sd_temp / sqrt(length(m_subset$Temperature))

# Calculate summary statistics for pH
m_mean_pH <- mean(m_subset$pH_T, na.rm=TRUE)
m_max_pH <- max(m_subset$pH_T, na.rm=TRUE)
m_min_pH <- min(m_subset$pH_T, na.rm=TRUE) 
m_sd_pH <- sd(m_subset$pH_T, na.rm=TRUE)
m_se_pH <- m_sd_pH / sqrt(length(m_subset$pH_T))

# Combine summary statistics into a table
(m_subset_summary_table <- data.frame(
  Variable = c("Temperature", "pH"),
  Mean = c(m_mean_temp, m_mean_pH),
  Minimum = c(m_min_temp, m_min_pH),
  Maximum = c(m_max_temp, m_max_pH),
  SD = c(m_sd_temp, m_sd_pH),
  SE = c(m_se_temp, m_se_pH)
))
```


### Checking if temperature is 25+/-10 in the mangrove
This is checked in dataframe m_subset where I subsetted the data from 10-02-2022 until 10-02-2023. We tested the accuracy of the HOBO Onset logger in a laboratory setting where we gradually increased temperature and measured the pH. After about 33/34 degrees Celcius, pH starts to decrease quite rapidly. We decided to removed measurements 25+/-10 as we are not sure about the accuracy of the data 10 degrees above and below the calibration temperature (following the manufacturer's statement).

```{r}
# Create an empty dataframe to store the rows that meet the condition
m_temp_table <- data.frame()

# For-loop through each row of the dataset
for (i in 1:nrow(m_subset)) {
  # check if the temperature is within the range
  if (m_subset$Temperature[i] >= 25 + 10 | m_subset$Temperature[i] <= 25 - 10) {
    # if the condition is true, add the row to the table
    m_temp_table <- rbind(m_temp_table, m_subset[i,])
  }
}

m_temp_table

# Remove rows in m_temp_table from the dataset m_subset and create new dataset
m_subset_removed <- m_subset[!(m_subset$V1 %in% m_temp_table$V1),] # .x1. %in% .x2. tests whether the values in x1 are in x2. The ! operator reverses the result 
str(m_subset_removed)
```

### Calculate summary statistics for m_subset_removed
```{r}
# Calculate summary statistics for temperature
m_removed_mean_temp <- mean(m_subset_removed$Temperature, na.rm=TRUE)
m_removed_max_temp <- max(m_subset_removed$Temperature, na.rm=TRUE)
m_removed_min_temp <- min(m_subset_removed$Temperature, na.rm=TRUE)
m_removed_sd_temp <- sd(m_subset_removed$Temperature, na.rm=TRUE)
m_removed_se_temp <- m_removed_sd_temp / sqrt(length(m_subset_removed$Temperature))

# Calculate summary statistics for pH
m_removed_mean_pH <- mean(m_subset_removed$pH_T, na.rm=TRUE)
m_removed_max_pH <- max(m_subset_removed$pH_T, na.rm=TRUE)
m_removed_min_pH <- min(m_subset_removed$pH_T, na.rm=TRUE) 
m_removed_sd_pH <- sd(m_subset_removed$pH_T, na.rm=TRUE)
m_removed_se_pH <- m_removed_sd_pH / sqrt(length(m_subset_removed$pH_T))

# Combine summary statistics into a table
(m_subset_removed_summary_table <- data.frame(
  Variable = c("Temperature", "pH"),
  Mean = c(m_removed_mean_temp, m_removed_mean_pH),
  Minimum = c(m_removed_min_temp, m_removed_min_pH),
  Maximum = c(m_removed_max_temp, m_removed_max_pH),
  SD = c(m_removed_sd_temp, m_removed_sd_pH),
  SE = c(m_removed_se_temp, m_removed_se_pH)
))

# See at which date/time the lowest pH value was recorded
(min_row_data <- m_subset_removed[min_row,])
```

### Visualization mangrove
```{r}
### All data (with data points outside 25 +/- 10 included)
ylab <- expression("Temperature " ( degree*C))
(m_temp <- ggplot(m_subset, aes(x=Date_Time, y=Temperature)) +
  geom_line(color="#33A02C") + 
  labs(x="", y=ylab, title="Inner mangrove") +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") + #%y for 22 %Y for 2022
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(14, 45), breaks = seq(15, 45, by = 5)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 10:16:00", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#33A02C"))

ylab <- "pH"
(m_pH <- ggplot(m_subset, aes(x=Date_Time, y=pH_T)) +
  geom_line(color="#33A02C") + 
  labs(x="", y=ylab) +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(3.5, 9), breaks = seq(3, 9, by = 1)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 10:16:00", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#33A02C"))

### Removed data (with data points outside 25 +/- 10 removed)
ylab <- expression("Temperature " ( degree*C))
(m_temp_removed <- ggplot(m_subset_removed, aes(x=Date_Time, y=Temperature)) +
  geom_line(color="#33A02C") + 
  labs(x="", y=ylab, title="Inner mangrove") +
  labs(x="", y=ylab) +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(size=10, angle=45, hjust=1), 
        axis.text.y=element_text(size=10),
        axis.title=element_text(size=12)) +
  scale_y_continuous(limits = c(14, 37.5), breaks = seq(15, 37.5, by = 5)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 10:16:00", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#33A02C"))

ylab <- "pH"
(m_pH_removed <- ggplot(m_subset_removed, aes(x=Date_Time, y=pH_T)) +
  geom_line(color="#33A02C") + 
  labs(x="", y=ylab) +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(size=10, angle=45, hjust=1), 
        axis.text.y=element_text(size=10),
        axis.title=element_text(size=12)) +
  scale_y_continuous(limits = c(5.5, 9), breaks = seq(5.5, 9, by = 1)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 10:16:00", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#33A02C"))
```


## Reef data 
```{r}
r_log <- read.csv("../2022-2023 Reef HOBO_Temp_pH.csv", header=FALSE, comment.char="#", stringsAsFactors=TRUE)

str(r_log)

# Rename multiple columns using rename()
# I had a conflict with another package so that rename() didn't work, use dplyr::rename and it worked (here you specifiy which package to use rename from)
r_log <- r_log %>% 
        dplyr::rename("Date_Time" = "V2",
               "Temperature" = "V3",
               "mV" = "V4",
               "pH_NBS" = "V5",
               "pH_T" = "V6")

# The time variable isn't at the date format, so time series data won't work. Use "lubridate" to convert it 
library(lubridate)
r_log$Date_Time <- as.POSIXct(r_log$Date_Time, format = "%d/%m/%Y %H:%M") 
str(r_log)

# Check for NA values in the dataset
anyNA(r_log) #no NAs in m_log
which(is.na(r_log$Date_Time)) #NA for date_time 2022-10-02 02:23:51
r_log$Date_Time[5720] <- as.POSIXct("2022-10-02 02:23:51")
anyNA(r_log) # no NAs anymore

# Set the range of datetimes (1 year)
r_start_time <- as.POSIXct("2022-02-10 0:23:51", format = "%Y-%m-%d %H:%M:%S")
r_end_time <- as.POSIXct("2023-02-10 23:23:51", format = "%Y-%m-%d %H:%M:%S")

# Subset the data within the datetime range
r_subset <- r_log[r_log$Date_Time >= r_start_time & r_log$Date_Time <= r_end_time, ]
anyNA(r_subset) # no NAs in m_subset

# Calculate summary statistics for temperature
r_mean_temp <- mean(r_subset$Temperature, na.rm=TRUE)
r_max_temp <- max(r_subset$Temperature, na.rm=TRUE)
r_min_temp <- min(r_subset$Temperature, na.rm=TRUE)
r_sd_temp <- sd(r_subset$Temperature, na.rm=TRUE)
r_se_temp <- r_sd_temp / sqrt(length(r_subset$Temperature))

# Calculate summary statistics for pH
r_mean_pH <- mean(r_subset$pH_T, na.rm=TRUE)
r_max_pH <- max(r_subset$pH_T, na.rm=TRUE)
r_min_pH <- min(r_subset$pH_T, na.rm=TRUE)
r_sd_pH <- sd(r_subset$pH_T, na.rm=TRUE)
r_se_pH <- r_sd_pH / sqrt(length(r_subset$pH_T))

# Combine summary statistics into a table
(r_subset_summary_table <- data.frame(
  Variable = c("Temperature", "pH"),
  Mean = c(r_mean_temp, r_mean_pH),
  Minimum = c(r_min_temp, r_min_pH),
  Maximum = c(r_max_temp, r_max_pH),
  SD = c(r_sd_temp, r_sd_pH),
  SE = c(r_se_temp, r_se_pH)
))
```


### Checking if temperature is 25+/-10 in the reef
This is checked in dataframe r_subset where I subsetted the data from 10-02-2022 until 10-02-2023. First few days are omitted from the dataset as logger needs to equilibrate first few days.

```{r}
# Create an empty dataframe to store the rows that meet the condition
r_temp_table <- data.frame()

# For-loop through each row of the dataset
for (i in 1:nrow(r_subset)) {
  # check if the temperature is within the range
  if (r_subset$Temperature[i] >= 25 + 10 | r_subset$Temperature[i] <= 25 - 10) {
    # if the condition is true, add the row to the table
    r_temp_table <- rbind(r_temp_table, r_subset[i,])
  }
}

r_temp_table
```

No datapoints fall outside the calibration range. So we can assume the pH values are all accurate.

### Visualization reef
Make 4 panel overview temp mangrove/reef A/B and pH mangrove/reef C/D, nrow=2
```{r}
ylab <- expression("Temperature " ( degree*C))
(r_temp_yrange <- ggplot(r_subset, aes(x=Date_Time, y=Temperature)) +
  geom_line(color="#1F78B4") + 
  labs(x="", y=ylab, title="Outer reef") +
  labs(x="", y=ylab) +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(size=10, angle=45, hjust=1), 
        axis.text.y=element_text(size=10),
        axis.title=element_text(size=12)) +
  scale_y_continuous(limits = c(14, 37.5), breaks = seq(15, 37.5, by = 5)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#1F78B4"))

(r_temp <- ggplot(r_subset, aes(x=Date_Time, y=Temperature)) +
  geom_line(color="#1F78B4") + 
  labs(x="", y=ylab, title="Outer reef") +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(14, 45), breaks = seq(15, 45, by = 5)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#1F78B4"))

ylab <- "pH"
(r_pH_yrange <- ggplot(r_subset, aes(x=Date_Time, y=pH_T)) +
  geom_line(color="#1F78B4") + 
  labs(x="", y=ylab) +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(size=10, angle=45, hjust=1), 
        axis.text.y=element_text(size=10),
        axis.title=element_text(size=12)) +
  scale_y_continuous(limits = c(5.5, 9), breaks = seq(5.5, 9, by = 1)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#1F78B4"))

(r_pH <- ggplot(r_subset, aes(x=Date_Time, y=pH_T)) +
  geom_line(color="#1F78B4") + 
  labs(x="", y=ylab) +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(3.5, 9), breaks = seq(3, 9, by = 1)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#1F78B4"))

#ggarrange(r_temp, r_pH)
```


### Shared plot mangrove & reef
Note: The first graph is made based on dataset with removed data points from the mangrove dataset that fell outside calibration range. Second graph is displayed in Supplementary Material with the full data set.

```{r}
library(ggpubr)
# Figure 1D
ggarrange(r_temp_yrange, m_temp_removed, r_pH_yrange, m_pH_removed, 
          #labels=c("A", "B", "C", "D"),
          ncol=2, nrow=2)
#ggsave("R_M_Year_resized.pdf", width=9, height=5)


# Supplementary Figure 3
ggarrange(r_temp, m_temp, r_pH, m_pH, 
          labels=c("A", "B", "C", "D"))
#ggsave("R_M_Year_all.pdf", width=9, height=5)
```


## Data from second set of loggers
### Mangrove 2
```{r}
m_log2 <- read.csv("../2022 Mangrove 2nd logger.csv", header=FALSE, comment.char="#")

# The time variable isn't at the date format, so time series data won't work. Use "lubridate" to convert it 
m_log2$Date_Time <- as.POSIXct(m_log2$Date_Time, format = "%d/%m/%Y %H:%M") #specify format m/d/Y
str(m_log2)

# Set the range of datetimes (mangrove logger stopped on 2022-10-22)
m_start_time2 <- as.POSIXct("2022-02-10 00:36:00", format = "%Y-%m-%d %H:%M:%S")
m_end_time2 <- as.POSIXct("2022-10-22 07:36:00", format = "%Y-%m-%d %H:%M:%S")

# Subset the data within the datetime range
m_subset2 <- m_log2[m_log2$Date_Time >= m_start_time2 & m_log2$Date_Time <= m_end_time2, ]

# Calculate the maximum and minimum values of the subset
(m_mean_temp2 <- mean(m_subset2$Temperature, na.rm=TRUE))
(m_max_temp2 <- max(m_subset2$Temperature, na.rm=TRUE)) 
(m_min_temp2 <- min(m_subset2$Temperature, na.rm=TRUE))

# Calculate summary statistics for temperature
m_subset2_mean_temp <- mean(m_subset2$Temperature, na.rm = TRUE)
m_subset2_max_temp <- max(m_subset2$Temperature, na.rm = TRUE)
m_subset2_min_temp <- min(m_subset2$Temperature, na.rm = TRUE)
m_subset2_sd_temp <- sd(m_subset2$Temperature, na.rm = TRUE)
m_subset2_se_temp <- m_subset2_sd_temp / sqrt(length(m_subset2$Temperature))

# Combine summary statistics into a table
m_subset2_summary_table <- data.frame(
  Variable = c("Temperature"),
  Mean = c(m_subset2_mean_temp),
  Minimum = c(m_subset2_min_temp),
  Maximum = c(m_subset2_max_temp),
  SD = c(m_subset2_sd_temp),
  SE = c(m_subset2_se_temp)
)

# Print the summary table
m_subset2_summary_table
```

```{r}
ylab <- expression("Temperature " ( degree*C))
(m_temp2 <- ggplot(m_subset2, aes(x=Date_Time, y=Temperature)) +
  geom_line(color="#33A02C") + 
  labs(x="", y=ylab, title="Mangrove") +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(14, 45), breaks = seq(15, 45, by = 5)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#33A02C"))

ylab <- expression("Temperature " ( degree*C))

(com_m <- ggplot() +
  geom_line(data=m_subset, aes(x=Date_Time, y=Temperature), color="#33A02C") +
  geom_line(data=m_subset2, aes(x=Date_Time, y=Temperature), color="grey", linetype="dashed") +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(14, 43), breaks = seq(15, 40, by = 5)) +
  labs(x="", y=ylab, title="Comparison 2 mangrove loggers") +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#33A02C"))

#ggsave(com_m)
```

### Reef 2
```{r}
r_log2 <- read.csv("../2022-2023 Reef 2nd logger.csv", header=FALSE, comment.char="#")

# The time variable isn't at the date format, so time series data won't work. Use "lubridate" to convert it 
r_log2$Date_Time <- as.POSIXct(r_log2$Date_Time, format = "%d/%m/%Y %H:%M") #specify format m/d/Y
str(r_log2)

# Set the range of datetimes
r_start_time2 <- as.POSIXct("2022-02-10 00:35:00", format = "%Y-%m-%d %H:%M:%S")
r_end_time2 <- as.POSIXct("2023-02-10 23:35:00", format = "%Y-%m-%d %H:%M:%S")

# Subset the data within the datetime range
r_subset2 <- r_log2[r_log2$Date_Time >= r_start_time2 & r_log2$Date_Time <= r_end_time2, ]

# Calculate the maximum and minimum values of the subset
(r_mean_temp2 <- mean(r_subset2$Temperature, na.rm=TRUE))
(r_max_temp2 <- max(r_subset2$Temperature, na.rm=TRUE)) 
(r_min_temp2 <- min(r_subset2$Temperature, na.rm=TRUE))

# Calculate summary statistics for temperature
r_subset2_mean_temp <- mean(r_subset2$Temperature, na.rm = TRUE)
r_subset2_max_temp <- max(r_subset2$Temperature, na.rm = TRUE)
r_subset2_min_temp <- min(r_subset2$Temperature, na.rm = TRUE)
r_subset2_sd_temp <- sd(r_subset2$Temperature, na.rm = TRUE)
r_subset2_se_temp <- r_subset2_sd_temp / sqrt(length(r_subset2$Temperature))

# Combine summary statistics into a table
r_subset2_summary_table <- data.frame(
  Variable = c("Temperature"),
  Mean = c(r_subset2_mean_temp),
  Minimum = c(r_subset2_min_temp),
  Maximum = c(r_subset2_max_temp),
  SD = c(r_subset2_sd_temp),
  SE = c(r_subset2_se_temp)
)

# Print the summary table
r_subset2_summary_table
```


### Visualisation
```{r}
ylab <- expression("Temperature " ( degree*C))
(r_temp2 <- ggplot(r_subset2, aes(x=Date_Time, y=Temperature)) +
  geom_line(color="#1F78B4") + 
  labs(x="", y=ylab, title="Reef 2") +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(14, 45), breaks = seq(15, 45, by = 5)) +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#1F78B4"))

(com_r <- ggplot() +
  geom_line(data=r_subset, aes(x=Date_Time, y=Temperature), color="#1F78B4") +
  geom_line(data=r_subset2, aes(x=Date_Time, y=Temperature), color="grey",linetype="dashed") +
  scale_x_datetime(date_breaks="1 month", date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_continuous(limits = c(20.0, 32), breaks = seq(15, 40, by = 2.5)) +
  labs(x="", y=ylab, title="Comparison 2 reef loggers") +
  geom_vline(xintercept=as.numeric(as.POSIXct("2022-10-23 14:16:20", format = "%Y-%m-%d %H:%M:%S")), linetype=2, color="#1F78B4"))
```


```{r}
# Supplementary Figure 5
ggarrange(com_m, com_r, nrow=2)
ggsave("Combined loggers.pdf", width=10, height=5)
```


#### Daily averages for mangrove m_subset_removed
With temperature and pH removed 25+/- 10
```{r}
### m_subset_removed
m_subset_removed$Month <- format(m_subset_removed$Date_Time, format="%B") #%b abbreviated month, %B unabbreviated month
m_subset_removed$Year <- format(m_subset_removed$Date_Time, format="%Y")
m_subset_removed$Day <- format(m_subset_removed$Date_Time, format="%d")
head(m_subset_removed)

m_removed_Daily_Avg <- aggregate(cbind(m_subset_removed$Temperature, m_subset_removed$pH_T),
                         by=list(Year=m_subset_removed$Day, Month=m_subset_removed$Month),
                                 FUN=mean)
names(m_removed_Daily_Avg) <- c("Day", "Month", "Average_Temperature", "Average_pH_T")
str(m_removed_Daily_Avg)
#write.csv(m_removed_Daily_Avg, file = "m_removed_Daily_Avg.csv")

# Calculate summary statistics for temperature
m_removed_Daily_Avg_mean_temp <- mean(m_removed_Daily_Avg$Average_Temperature)
m_removed_Daily_Avg_max_temp <- max(m_removed_Daily_Avg$Average_Temperature, na.rm=TRUE)
m_removed_Daily_Avg_min_temp <- min(m_removed_Daily_Avg$Average_Temperature, na.rm=TRUE)
m_removed_Daily_Avg_sd_temp <- sd(m_removed_Daily_Avg$Average_Temperature, na.rm=TRUE)
m_removed_Daily_Avg_se_temp <- m_removed_Daily_Avg_sd_temp / sqrt(length(m_removed_Daily_Avg$Average_Temperature))

# Calculate summary statistics for pH
m_removed_Daily_Avg_mean_pH <- mean(m_removed_Daily_Avg$Average_pH_T, na.rm=TRUE)
m_removed_Daily_Avg_max_pH <- max(m_removed_Daily_Avg$Average_pH_T, na.rm=TRUE)
m_removed_Daily_Avg_min_pH <- min(m_removed_Daily_Avg$Average_pH_T, na.rm=TRUE) 
m_removed_Daily_Avg_sd_pH <- sd(m_removed_Daily_Avg$Average_pH_T, na.rm=TRUE)
m_removed_Daily_Avg_se_pH <- m_removed_Daily_Avg_sd_pH / sqrt(length(m_removed_Daily_Avg$Average_pH_T))

# Combine summary statistics into a table
(m_removed_Daily_Avg_summary_table <- data.frame(
  Variable = c("Temperature", "pH"),
  Mean = c(m_removed_Daily_Avg_mean_temp, m_removed_Daily_Avg_mean_pH),
  Minimum = c(m_removed_Daily_Avg_min_temp, m_removed_Daily_Avg_min_pH),
  Maximum = c(m_removed_Daily_Avg_max_temp, m_removed_Daily_Avg_max_pH),
  SD = c(m_removed_Daily_Avg_sd_temp, m_removed_Daily_Avg_sd_pH),
  SE = c(m_removed_Daily_Avg_se_temp, m_removed_Daily_Avg_se_pH)
))
```


#### Daily averages for reef r_subset 
```{r}
### Reef daily averages
r_subset$Month <- format(r_subset$Date_Time, format="%B") #%b abbreviated month, %B unabbreviated month
r_subset$Year <- format(r_subset$Date_Time, format="%Y")
r_subset$Day <- format(r_subset$Date_Time, format="%d")
head(r_subset)

r_Daily_Avg <- aggregate(cbind(r_subset$Temperature, r_subset$pH_T),
                         by=list(Year=r_subset$Day, Month=r_subset$Month),
                         FUN=mean)
names(r_Daily_Avg) <- c("Day", "Month", "Average_Temperature", "Average_pH_T")
str(r_Daily_Avg)
#write.csv(r_Daily_Avg, file = "r_Daily_Avg.csv")

# Calculate summary statistics for temperature
r_Daily_Avg_mean_temp <- mean(r_Daily_Avg$Average_Temperature)
r_Daily_Avg_max_temp <- max(r_Daily_Avg$Average_Temperature, na.rm=TRUE)
r_Daily_Avg_min_temp <- min(r_Daily_Avg$Average_Temperature, na.rm=TRUE)
r_Daily_Avg_sd_temp <- sd(r_Daily_Avg$Average_Temperature, na.rm=TRUE)
r_Daily_Avg_se_temp <- r_Daily_Avg_sd_temp / sqrt(length(r_Daily_Avg$Average_Temperature))

# Calculate summary statistics for pH
r_Daily_Avg_mean_pH <- mean(r_Daily_Avg$Average_pH_T, na.rm=TRUE)
r_Daily_Avg_max_pH <- max(r_Daily_Avg$Average_pH_T, na.rm=TRUE)
r_Daily_Avg_min_pH <- min(r_Daily_Avg$Average_pH_T, na.rm=TRUE) 
r_Daily_Avg_sd_pH <- sd(r_Daily_Avg$Average_pH_T, na.rm=TRUE)
r_Daily_Avg_se_pH <- r_Daily_Avg_sd_pH / sqrt(length(r_Daily_Avg$Average_pH_T))

# Combine summary statistics into a table
(r_Daily_Avg_summary_table <- data.frame(
  Variable = c("Temperature", "pH"),
  Mean = c(r_Daily_Avg_mean_temp, r_Daily_Avg_mean_pH),
  Minimum = c(r_Daily_Avg_min_temp, r_Daily_Avg_min_pH),
  Maximum = c(r_Daily_Avg_max_temp, r_Daily_Avg_max_pH),
  SD = c(r_Daily_Avg_sd_temp, r_Daily_Avg_sd_pH),
  SE = c(r_Daily_Avg_se_temp, r_Daily_Avg_se_pH)
))
```


### Tables for mean/min/max with SD and SE for everything
```{r}
# With measurements every hour
m_subset_removed_summary_table
m_subset_summary_table
r_subset_summary_table

# Daily averages
m_removed_Daily_Avg_summary_table
#m_Daily_Avg_summary_table
r_Daily_Avg_summary_table
```

### Statistics Daily averages
```{r}
daily_av <- read.csv("../Daily_average_M_R.csv", stringsAsFactors=TRUE)

str(daily_av)
library(lubridate)
daily_av$Day_month <- as.POSIXct(paste0(daily_av$Day_month, "-2022"), format = "%d-%b-%Y")
str(daily_av)

hist(daily_av$Av_temp)

# Data transformations for negatively skewed/left-skewed data
daily_av$Av_temp_sqrt <- sqrt(max(daily_av$Av_temp+1) - daily_av$Av_temp)
daily_av$Av_temp_log <- log10(max(daily_av$Av_temp+1) - daily_av$Av_temp)
daily_av$Av_temp_inv <- 1/(max(daily_av$Av_temp+1) - daily_av$Av_temp)

#Shapiro test
shapiro.test(log(daily_av$Av_temp))
shapiro.test(daily_av$Av_temp_sqrt)
shapiro.test(daily_av$Av_temp_log)
shapiro.test(daily_av$Av_temp_inv)
####not passed#### 

# Create a boxplot of Av_temp by Habitat
boxplot(Av_temp ~ Habitat, data = daily_av, 
        xlab = "Habitat", ylab = "Av_temp", 
        main = "Boxplot of Av_temp by Habitat")


library(car)
#Levene test
leveneTest(Av_temp ~ Habitat, data=daily_av)
####passed####

wilcox.test(Av_temp ~ Habitat, data = daily_av)

# Perform two-sample t-test on Av_temp by Habitat
#t.test(Av_temp ~ Habitat, data = daily_av)

#Shapiro test
shapiro.test(daily_av$Av_pH_T)
####not passed#### 

#Levene test
leveneTest(Av_pH_T ~ Habitat, data=daily_av)
####not passed####

# Perform two-sample t-test on Av_pH by Habitat
#t.test(Av_pH_T ~ Habitat, data = daily_av)
wilcox.test(Av_pH_T ~ Habitat, data = daily_av)
```


### Statistics Daily maximum
```{r}
r_subset$Time <- format(r_subset$Date_Time, "%H:%M:%S")
r_subset_15 <- subset(r_subset, Time == "15:23:00")
#r_max_temp <- aggregate(Temperature ~ as.Date(Date_Time), r_subset_15, max)
#write.csv(r_subset_15, file = "r_Max_15.csv")

m_subset_removed$Time <- format(m_subset_removed$Date_Time, "%H:%M:%S")
m_subset_removed_15 <- subset(m_subset_removed, Time == "15:16:00")
#m_subset_removed_max_temp <- aggregate(Temperature ~ as.Date(Date_Time), m_subset_removed_15, max)
#write.csv(m_subset_removed_15, file = "m_removed_Max_15.csv")

daily_max <- read.csv("../Daily_max_M_R.csv", stringsAsFactors=TRUE)
daily_max$Date_Time <- as.POSIXct(daily_max$Date_Time, format = "%d/%m/%Y %H:%M")
str(daily_max)

# Mean peak temperature and SEs
# Calculate mean temperature for each habitat
mean_temperatures <- aggregate(daily_max$Temperature, by = list(daily_max$Habitat), mean)

# Calculate standard errors for each habitat using SEM formula
sem_temperatures <- tapply(daily_max$Temperature, daily_max$Habitat, function(x) sd(x)/sqrt(length(x)))

# Combine mean and standard error into a data frame
habitat_summary <- data.frame(Habitat = mean_temperatures$Group.1,
                              Mean_Temperature = mean_temperatures$x,
                              Standard_Error = sem_temperatures)

# Print the summary
print(habitat_summary)

# Quick visualisation
ggplot(daily_max, aes(y=Temperature, x=Date_Time, color=Habitat)) +
  geom_point() +
  geom_line()

ggplot(daily_max, aes(y=pH_T, x=Date_Time, color=Habitat)) +
  geom_point() +
  geom_line()

#Shapiro test
shapiro.test(daily_max$Temperature)
####not passed#### 

#Levene test
library(car)
leveneTest(Temperature ~ Habitat, data=daily_max)
####not passed####

# Perform two-sample t-test on max_temp by Habitat
#t.test(Temperature ~ Habitat, data = daily_max)
wilcox.test(Temperature ~ Habitat, data = daily_max)

# Perform two-sample t-test on max_pH_T by Habitat
#t.test(pH_T ~ Habitat, data = daily_max)
```